### 2022-04-22 paper/paper_draft.Rmd
more paper updates, reworked to not conclude which re-emergence hypothesis is correct and to focus on proximate vs facultative genes.

---
title: "Novel genetic control of migratory diapause in Australian Monarch Butterflies"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    reference_docx: "paper-files/style_ref.docx"
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    fig_width: 7
    fig_height: 6
bibliography: "paper-files/citations.bib"
csl: "paper-files/council-of-science-editors-author-date.csl"
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}\doublespacing
  - \usepackage{sectsty}\sectionfont{\fontsize{12}{12}\selectfont}
  - \usepackage{sectsty}\subsectionfont{\normalfont\itshape\fontsize{12}{12}\selectfont}
  - \usepackage[round]{natbib}
indent: true
sansfont: Times New Roman
fontsize: 12pt
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(snpR); library(data.table); library(ggplot2); library(gridExtra); library(dplyr); library(openxlsx); library(lme4)

```

<!-- Introduction: -->
<!--   Migration, partial migration, and adaptation to environmental change. -->
<!--   Monarchs in the Pacific and Australia, History -->
<!--   Two options for migratory behavior in Aus (proximate vs ultimate causes as well) -->
<!--   what we did -->

# Introduction

Migratory species are acutely at risk from global anthropogenic environmental change, since they rely on multiple distinct ecosystems and consistent seasonal timings [@Both2010]. As a result, the loss of breeding, wintering, or any transitional habitats may all cause declines in migrants [@Robbins1989], as may shifts in resource phenology due to seasonal changes which can cause resource availability and species life-histories to fall out of synch [@Jones2010a]. It is therefore not surprising that widespread declines have been observed in many migratory taxa [@Limburg2009;@Robbins1989].

Population-level adaptability in migratory life-history should therefore be expected to increase long-term population viability by conferring resistance to phenological and environmental changes [@Pulido7341]. Migratory life-history adaptation can essentially occur along three distinct axis: 1) *temporal*, or adaptation that changes the phenology of migration; 2) *spatial*, or adaptation that alters migratory pathways and orientation; and 3) *residency*, or adaptation that shifts populations between migratory and resident life histories. Inter-population variation and adaptation in each of these pathways is well known from nature: temporal changes or variation in migration are well documented in many taxa, particularly in birds [@Rubolini2007; @Gordo2007]; short-stopping, extension of migratory pathways, or complete re-direction of migration is well known in birds [@Ambrosini2011; @LaSorte2007; @Visser2009]; and shifts towards residency or partial migration have been observed in birds [@Pulido7341; @Plummer2015; @Møller2014], butterflies [@Howard2010], mammals [@VanDerRee2006; @Fleming2003; @Jones2014a], and other taxa.

Adaptation via a transition to residency is of both particular interest and concern for the conservation of migratory species. While switching to a resident life history can allow species to avoid the increasing risks of migration [and the intrinsically very energetic large costs of migration in general, see @McWilliams2004; @Rankin1992; @Wikelski2003], there is reasonable concern that the "extinction" of migration may not be easily reversible in some species. This is concerning since migratory life-history variation is expected to increase population adaptability [@Pulido2007], and thus the permanent loss of migration could potentially compromise the long term population persistence.

This is, for example, a major concern for Monarch Butterflies (Danaus plexippus). North American migratory monarch populations have experienced severe declines over the last several decades, sparking a fear of what Semmens et al. [-@Semmens2016] called a "Quasi-extinction" of the migratory phenotype driven by changes in the selective forces acting on the species to favor resident individuals. In addition to the damage to the long-term population persistence that could result from the loss of migration in North American, the annual monarch migration is publicly-beloved and its loss would be culturally costly [@Gustafsson2015].

Interestingly, however, there is at least one population of Monarchs in Australia in which migratory behavior has re-emerged after previously shifting to a resident life history. This population, which was established sometime before 1871, is the product of a series of sequential self-introductions that stepping-stoned across the Pacific ocean from North America beginning sometime in the early 1840s after the introduction of their host plants enabled individuals to survive after being blown from the mainland [@Zalucki2004]. Monarchs are not migratory on any Pacific islands, and no migratory behavior was observed until the 1930s, when non-reproductive overwintering aggregations were observed in northern New South Wales [NSW, see @James2019]. These and other individuals were later shown to migrate hundreds of kilometers each year to their wintering grounds from more southerly, inland locations in NSW [@James2019]. After over 50 years (or hundreds of generations) without migration, a migratory life-history re-emerged in these populations. Understanding how this happened could provide crucial context around the "extinction" of migration in monarchs and other species.

To better understand how migration re-emerged in Australian monarchs, we studied continuously breeding monarch populations in Queensland (QLD). While these populations do not experience substantial seasonal fluctuations in resources availability, and thus do not migrate [@James2019], recent work has shown that reproductive diapause can be induced in some, *but not all* individuals from these populations if they are exposed to a reduction in photoperiod during development [@Freedman2018]. In brief, reproductive diapause, where individuals delay their reproductive maturation, is a critical phenotype for migratory individuals because it allows them to invest more heavily in energy storage and increase their average longevity [@Herman2001]. Like other migratory behaviors, diapause is controlled by changes in day length, temperature, and other seasonal fluctuations [@Goehring2002].

The variance in diapause induction in QLD monarchs allows for a direct search for genetic associations with the trait and thus for an examination of the genetic causes of their re-evolution of migration. *Here, we used a mix of reduced representation and whole genome sequencing to search for genes associated with migratory diapause in QLD monarch butterflies. We found one strongly associated locus which had previously been identified to be associated with juvenile phase transitions in silk moths (insert species name), which is controlled by the same major hormone as reproductive diapause. This gene had not previously been found to be associated with migration in North American monarch butterflies, suggesting that diapause and potentially migration as a whole was secondarily re-evolved in Australian monarchs by novel mutations.*

# Methods
## Sample Collection, Incubation, and Phenotyping
```{r, read_sample_meta_get_numbers, echo=FALSE,message=FALSE,include=FALSE,warning=FALSE}
sample.meta <- read.table("../data/sample_metadata.txt", header = T, stringsAsFactors = F)
old.data <- data.table::fread("../data/Aus_2018_metadata.txt", header = T)
old.data[,fam := as.numeric(gsub("_.+", "", ID))]

meta.micah <- data.table::fread("../data/2016_mon_meta.txt", header = TRUE)
```

We collected 22 female monarch butterflies from Pinjarra Hills, Queensland, Australia (27°32′26.7″S, 152°54′22.7″E) between the 5th and 9th of July 2018. We noted individuals in all life stages present at the field site during our sample collections. We then enclosed these individuals on their host plants and collected the eggs produced by 21 of them. These eggs represent 21 total material families, but, since females may mate multiple times and lay clutches with mixed parentage [CITE], each of these families may consist of multiple groups of half-siblings. We did not remove *Ophryocystis elektroscirrha* (a protozoan parasite) spores, which were uniformly present on our eggs [consistent with previous reports from this location, see @Freedman2018].

<!-- add citation -->

We then incubated all of the eggs using the same incubators and following the "decreasing photoperiod" experimental scheme of Freedman et al. [-@Freedman2018] in order to induce reproductive diapause and ensure that our data could be later pooled with that produced in the earlier study. We released all male butterflies following emergence, since phenotyping diapause in male monarchs can be challenging due to the difficulty in extracting and weighing the seminal vesicle, which is likely the best indicator of male reproductive development [@Freedman2018]. This left us with a total of `r nrow(old.data)` female individuals from `r length(unique(na.omit(old.data$fam)))` maternal families.

We then assessed reproductive diapause by assessing the degree of oocyte development in each of our adult female butterflies. We observed that individuals fell into two clearly defined bins: `r sum(old.data$Phenotype == "C", na.rm = TRUE)` out of `r nrow(old.data)` had fully developed, chorionated eggs (visible ridges along the exterior of the egg), `r sum(old.data$Phenotype != "C", na.rm = TRUE)` females did not. Of these latter individuals a few (`r sum(old.data$Phenotype == "I", na.rm = TRUE) + sum(old.data$Phenotype == "N", na.rm = TRUE)`) had no yolk in their eggs whatsoever, but most had some degree of yolking. Since vitellogenesis generally only begins to a substantial degree after eclosion in monarchs [@Pan1976], We called any individuals with more than a small amount of yolking in their eggs as reproductively mature (N = `r sum(sample.meta$likely_pheno[sample.meta$year == 2018] == "M", na.rm = TRUE)` and `r sum(sample.meta$likely_pheno[sample.meta$year == 2018] == "R", na.rm = TRUE)` for reproductively immature and mature individuals, respectively, see Supplementary Information for images of egg development in each individual). Wings and bodies from all individuals were preserved in dry coin envelops or 95% ethanol, respectively.

We supplemented these samples with preserved, dry butterfly wings from each of the "decreasing photoperiod" female individuals previously phenotyped for reproductive development by Freedman et al. [-@Freedman2018], constituting an additional `r sum(sample.meta$year == 2016)` total female butterflies from `r length(unique(meta.micah$Female_Line[meta.micah$Treatment == "Decreasing"]))` maternal families, of which `r sum(sample.meta[sample.meta$year == 2016,]$likely_pheno == "M", na.rm = TRUE)` and `r sum(sample.meta[sample.meta$year == 2016,]$likely_pheno == "R", na.rm = TRUE)` were phenotyped by the authors as reproductively immature and mature, respectively. Together, these two datasets contained `r length(na.omit(sample.meta$likely_pheno))` monarch butterflies, of which `r sum(sample.meta$likely_pheno == "M", na.rm = TRUE)` and `r sum(sample.meta$likely_pheno == "R", na.rm = TRUE)` were reproductively immature and mature, respectively (see Table 1).

## Sequencing and Genotyping
We removed and extracted DNA one leg from each of our samples collected in 2018 and a wing base from each of the samples collected by Freedman et al [-@Freedman2018] using the magnetic bead protocol of Ali et al [-@Ali389]. We quantified the resulting DNA on a BioTek Instruments FLx800 Fluorescence Reader using Thermo Fisher Scientific Quant-iT PicoGreen dsDNA Reagent, the prepared Restriction Associated Digest (RAD) libraries using the Pst1 restriction enzyme according to Ali et al [-@Ali389]. We sequenced these 150 bp paired-end sequencing libraries using an Illumina Hi-Seq 4000.

We aligned the resulting raw sequence data to the v4 monarch butterfly genome assembly [@Gu2019] using the mem algorithm of the Burrows-Wheeler Aligner [@Li2009a]. We then filtered out PCR duplicates, improperly paired, or poorly mapped reads using SAMtools [@Li2009].

<!-- we removed poorly seqeuenced individuals later in the pipeline, so these numbers are actually the final numbers. Not sure if that matters, but can fix. -->

<!-- and then used the ngsParalog tool to removed an potentially paralogous loci by restricting our analysis to loci more than a kilobase away from loci with a log ratio test statistic > 10 [@Linderoth2018]. -->

<!-- Didn't do paralog, redoing now did do a rough HWE filter. Probably shouldn't really do paralog check, though, since there is a lot of family structure in our samples, which will throw things off. -->

We then called gentoypes using the ANGSD software package with the following parameters: -doMajorMinor 1 (determine major and minor alleles using a genotype likelihood approach),  -doMaf 2 (determine minor allele frequencies), -SNP_pval 1e-8 (keep only loci with a SNP p-value $\le 1 \times 10^{-8}$), -doGeno 4 (call genotypes), -doPost 2 (calculate genotype posterior probabilities using a uniform prior), -postCutoff 0.95 (keep only loci where the highest genotype posterior probability $ \ge 0.95$), -minQ 20 (keep only loci with a sequencing quality $\ge 20$), -minMapQ 20 (keep only loci with a mapping quality $\ge 20$), -minInd 130 (keep only loci sequenced in at least 130 individuals), and -minMaf 0.05 (keep only loci with a minor allele frequency $\ge 0.05$).

## Statistical Analysis

```{r import_and_diversity, echo=FALSE, message=FALSE, warning=FALSE,include=FALSE}
# L <- # total number of sequenced bases

#==========prep data==========
# read in genotypes and metadata
sample.meta <- read.table("../data/sample_metadata.txt", header = T, stringsAsFactors = F)
genos <- read.table("../data/v4/genotypes.geno", stringsAsFactors = F)
snp.meta <- genos[,1:2]
genos <- genos[,-c(1:2)]
colnames(snp.meta) <- c("scaffold", "position")

# rename msc scaffolds
chrs <- which(grepl("chr",snp.meta$scaffold))
snp.meta$scaffold[-chrs] <- "other"
snp.meta$scaffold <- factor(snp.meta$scaffold, levels = c(paste0("chr", 1:30), "other"))

# remove samples with no phenotypic data
nas <- which(is.na(sample.meta$likely_pheno))
genos <- genos[,-which(is.na(sample.meta$likely_pheno))]
sample.meta <- sample.meta[-which(is.na(sample.meta$relaxed_pheno)),]

# add in colony info
fam_info <- sample.meta(readRDS("../colony/colony_R_outputs.RDS")$col_res$x)
fam_info$MotherID <- as.numeric(gsub("#", "", fam_info$MotherID))
fam_info$MotherID <- factor(fam_info$MotherID, levels = 1:max(fam_info$MotherID))
sample.meta$MotherID <- fam_info$MotherID # note, individuals are in identical order.

# send year to factor, import, and filter
sample.meta$year <- as.factor(sample.meta$year)
# dat <- import.snpR.data(genos, snp.meta, sample.meta)
# # dat <- dat[,-which(sample.meta(dat)$year != 2018)] # note: same locus pops out if you only use my samples, but is stronger with both. Good sign!
# dat <- filter_snps(dat, hwe = 0.000001, hwe_facets = "year")

#==========diversity===========
# dat <- calc_pi(dat, c(".base", "year"))
# dat <- calc_ho(dat, c(".base", "year"))
# dat <- calc_fis(dat, c(".base", "year"))
# dat <- calc_pairwise_fst(dat, "year", boot = 100, boot_par = 6, verbose = TRUE)
# saveRDS(dat, "diversity_stats.RDS")
dat <- readRDS("diversity_stats.RDS")
diversity <- get.snpR.stats(dat, c(".base", "year"), c("pi", "ho", "fis"))$weighted.means
fst <- get.snpR.stats(dat, "year", "fst")$weighted.means

#==========PCA===============
pca <- plot_clusters(dat, c("year.MotherID")) # color by mother and by year

```

In order to determine if any genomic regions were significantly associated with We used the snpR R package to filter out loci significantly out of Hardy-Weinburg Equilibrium ($p \le 0.000001$) in either year in order to help remove any paralogous loci and interpolated missing data by conducting binomial draws at missing genotypes using the minor allele frequencies at those loci [@Hemstrom2021]. To assess the basic level of genetic diversity amongst our samples, we then calculated the average number of pairwise differences ($\pi$), observed heterozygosity ($H_{O}$), and $F_{IS}$ according to Weir and Cockerham [-@Weir1984] for all of our samples and for each year independently using snpR [@Hemstrom2021]. We also calculated $F_{ST}$ according to Weir and Cockerham [-@Weir1984] between sample years and between both sample year and reproductive phenotype jointly[@Hemstrom2021]. We bootstrapped a significance level for $F_{ST}$ by bootstrapping individuals among populations with replacement 100 times. To determine if there was any substantial unexpected genetic structuring (other than family groups) in our samples, we additionally both conducted a Principal Component Analysis using snpR [@Hemstrom2021]. Since some individuals escaped their rearing dishes (but not the incubator) during rearing, we also confirmed the parentage of our samples using the Colony2 program, assuming polygamous, random mating [@Jones2010].

```{r association, echo=FALSE, warning=FALSE, message=FALSE}
# family bias
bias <- sample.meta[,c("MotherID", "likely_pheno")]
bias$MotherID <- as.factor(bias$MotherID)
bias$likely_pheno <- as.factor(bias$likely_pheno)
fisher_res <- fisher.test(bias$MotherID, bias$likely_pheno, simulate.p.value = TRUE)

bias_p <- reshape2::dcast(bias, MotherID ~ likely_pheno)
bias_p$total <- bias_p$M + bias_p$R
bias_p$proportion <- bias_p$M/bias_p$total
# Figure2a <- ggplot(bias, aes(x = bias$MotherID, fill = bias$likely_pheno)) + geom_bar(position = 'dodge')
Figure2a <- ggplot(bias_p, aes(x = MotherID, y = proportion, size = total)) + geom_point() + theme_bw() + geom_hline(yintercept = .5) + xlab("Maternal Family") + ylab("Proportion Diapaused") + labs(size = "n")
Figure2al <- ggpubr::get_legend(Figure2a)
Figure2a <- Figure2a + theme(legend.position = "none")

# association test
dat <- calc_association(dat, response = "likely_pheno", maxiter = 1000)

# plot
Figure2b <- plot_manhattan(dat, "gmmat_pval_likely_pheno", chr = "scaffold", significant = 0.00005, suggestive = 0.0005, log.p = T,
                           chr_order = c(paste0("chr", 1:30), "other"), highlight_style = "red") # highlight top snp
Figure2b <- Figure2b$plot + ylab("-log10(p-value)") + xlab("Position") + theme(strip.text = element_blank())
```
In order to determine the degree to which diapause status was biased between families, we constructed a Fisher's exact test for independence using 2000 simulations in R. We then conducted a Genome-Wide Association Study (GWAS) in order to determine if any loci were significantly associated with reproductive status. Since our samples were composed of many groups of full and half-siblings and came from two genetically different years, we used the GMMAT R package to conduct a population and family structure-corrected GWAS [@Chen2020; @Chen2016]. We visualized the results by constructing a Manhattan and quantile-quantile plot using snpR [@Hemstrom2021]. We identified the genes underlying any candidate regions by hand using the MonarchBase genome browser online tool [@Zhan2013].

```{r re_analyze, include=FALSE}
old_phenotypic <- read.xlsx("../data/monarch_development.xlsx")

mod_AsRa <- lme(Aspect_Ratio ~ Yolked_Oocytes, random = ~ 1|Female_Line, old_phenotypic, na.action = na.omit)
mod_Area <- lme(Marea ~ Yolked_Oocytes, random = ~ 1|Female_Line, old_phenotypic, na.action = na.omit)

```

We also briefly re-analyzed phenotypic data from the same earlier study that provided our 2016 monarch studies [@Freedman2018] in order to determine if wing shape or size was correlated with reproductive diapause by fitting a linear mixed effect model with wing shape or size as response variables against the number of yolked oocytes as a fixed effect and maternal family as a random effect using the R package nnlme [@Pinheiro2007].

# Results
## Sequencing results
We sequenced a total of 417,846,352 reads across all individuals, 98.8% of which mapped to the monarch reference genome. 244,443,414 reads were maintained after filtering (1,158,500 per sample, on average), encompassing X total sequenced bases. From this, we called `r prettyNum(nsnps(dat),big.mark=",",scientific=FALSE)` individual SNPs following filtering.

## Diversity and Structuring

Genetic diversity levels were extremely similar between our two years ($\pi =$ `r round(diversity$weighted_mean_pi[2], 3)` and `r diversity$weighted_mean_pi[3]` and $H_{o}$ = `r round(diversity$weighted_mean_ho[2], 3)` and `r round(diversity$weighted_mean_ho[3], 3)` for 2016 and 2018, respectively). Each year had a slight heterozygote excess ($F_{IS}$ = `r round(diversity$weighted_mean_fis[2], 3)` and `r round(diversity$weighted_mean_fis[3],3)` for 2016 and 2018, respectively), although that excess was lower if all samples were considered jointly ($F_{IS}$ = `r round(diversity$weighted_mean_fis[1], 3)`). Pairwise $F_{ST}$ between the two years was small but marginally significant (`r round(fst$weighted_mean_fst[1], 3)`, $p =$ `r round(fst$weighted_mean_fst_p[1], 3)`. We found no significant differences between the two sample years on either the PCA or the NGSadmix results, with clustering instead driven by family (Figure 1).

## Genetic Basis of Diapause

Diapuse status was strongly biased by Maternal Family (Fisher's Exact Test, $p$ = `r round(fisher_res$p.value)`, see Figure 2a). We identified one small region of chromosome 11 roughly between positions 5,589,278 to 5,619,459 containing many loci associated with diapause status well beyond the background level of association, including three SNPs from two different RAD-tags that had a $p$-value  $\le 5 \times 10^{-5}$ (Figure 2b). This region overlaps the gene DPOGS208560, a protein modifying gene, and was still strongly associated with diapause if only the 2018 butterflies alone were examined. We also identified one locus with a $p$-value  $\le 5 \times 10^{-5}$ on chromosome 16, but there were no other significantly associated loci nearby.

We found that wing shape was significantly negatively correlated with the number of yolked oocytes in Freedman et al.'s [-@Freedman2018] earlier study ($p$ = `r round(summary(mod_AsRa)$tTable[2,5], 3)`) after accounting for maternal family. Wing size was also negatively correlated with the number of yolked oocytes, but not significantly so after accounting for maternal family ($p$ = `r round(summary(mod_Area)$tTable[2,5], 3)`.

# Discussion

<!-- Outline: -->
<!--   The gene we found and why it makes sense -->
<!--   The gene we found in the context of migratory genes elsewhere in monarchs -->
<!--       This means that this is likely a novel mutation, not conserved variation -->
<!--   What this means for the persistence of migration in monarchs -->
<!--   What this means for the persistence and adaptability of migration in general. -->

## Re-emergence of Diapaue Control in Australian Monarch Butterflies

We found that a region on Chromosome 11 containing the gene DPOGS208560 was strongly associated with diapuase initiation in Monarch Butterflies in QLD, Australia. Interestingly, a predicted homolog of this gene in silkworms (*Bombyx mori*) has been previously differential expressed during larval molt-intermolt transitions [@Hu2016], which are generally controlled by fluctuations in juvenile hormone (JH) and ecdysone [@Jindra2013]. JH is also known to be a major driver of reproductive development and diapause in monarchs [@Herman2001], so it is not particularly surprising that DPOGS208560 could contribute to diapause control in monarchs.

What is surprising, however, is that this gene has not been previously identified as a candidate migration gene in monarchs previously. Specifically, DPOGS208560 was not among the 536 genes located in regions of the genome found to be significantly associated with contrasts between migration and residency between North and South American, Pacific, and European populations by Zhan et al. [-@Zhan2014]. There are two distinct possibilities for that could explain the origin of this gene variant: the DPOGS208560 variant we observed is either present in North America but was not previously recognized as associated with migration or is the result of a new mutation which occurred in Australian monarchs. In the former case, this allele was either maintained as monarchs passed through repeated population bottlenecks during their Pacific expansion or was secondarily re-introduced by immigrant individuals from North America (which would have likely been anthropogenic in origin).

Since we *only* observed that only the DPOGS208560 gene was associated with diapause in Australian monarchs, it seems unlikely that secondary reintroduction of migratory monarch butterflies from North America could explain the re-emergence of migratory behavior in Australia, since there would be no reason for other *known* migratory associated genes to have not also made the jump back across the pacific. Since the SNP loci that we located are not perfectly associated with diapause and are thus likely to only be in linkage with any causal variation rather than the causal variants themselves, our work cannot currently delineate between the other two options (maintenance of variance at DPOGS208560 during the pacific crossing or a new mutation). If the actual causal loci were to be identified, it would then be possible to determine which gene versions are present in North America or across the Pacific and thus determine the origin of the DPOGS208560 diapause variant in Australia.

## Proximate vs Facultative Migratory Alleles

Regardless, however, if it is interesting that DPOGS208560 had not been previously noted as associated with migration in monarch butterflies. In their 2014 study, Zhan et al. generally focused their discussion of migratory genetics on Collagen IV $\alpha$-1, one of three genes they found that were strong outliers for association with migration. This gene is involved in muscle functioning in insects [@Schnorrer2010], and Zhan et al hypothesized that divergence at this gene was driven by selection for increased wing strength and muscle efficiency to facilitate long-distance migration. While they mentioned that the remaining significantly associated genes were enriched for the "morphogenesis, neurogenesis, and extracellular matrix/basement membrane" functional terms, they were not otherwise discussed (DPOGS208560 does not have any of these functional terms). Note that it is entirely possible that DPOGS208560 also contributes to diapause and migration in butterflies outside of Australia, but that the quantitative effect of other genes (a product of the effect sizes of the migratory alleles and their frequencies) dwarfs that of DPOGS208560, thus concealing the relative impact of the latter gene.

Regardless, as mentioned above, we did not identify loci near the Collagen IV $\alpha$-1 gene as associated with diapause. This is surprising, since although monarchs in Australia (like other non-migratory populations), have substantially smaller and shorter wings than their migratory counterparts [@Freedman2020], we also found from re-analyzing data from Freedman et al.'s 2018 study, wing shape is significantly correlated with reproductive development in Australian monarchs, which in implies that wing morphology is still a part of the correlated migratory syndrome in Australian monarchs.

However, it does make some sense that Collagen IV $\alpha$-1 would not be a key gene in populations that are *newly* migratory. We would argue that migratory genes can be divided into two basic categories: facultative migratory genes and those that are the proximate drivers of migratory behavior. Collagen IV $\alpha$-1 is, for example, a "facultative" migratory gene in that selection acting on it would serve to make already migratory individuals more fit, but would not itself *cause* or *enable* migratory behavior. DPOGS208560, on the other hand, is likely a proximate migratory gene, since diapause is *required* for migration because individuals that do not delay their reproductive investment have much shorter life-spans and cannot complete migration [@Herman2001]. Genes that control orientation, navigation, and directed flight would also fall into this category to varying extents. Any population that is newly migratory must have some proximate causal gene migratory variants or they would not display migratory behavior in the first place, but may only acquire facultative migratory gene variants over time as selection favors migratory individuals that are better able to complete their migration. That we found DPOGS208560 but not Collagen IV $\alpha$-1 to be associated with migration in Australia supports this hypothesis, and generally supports a complex interpretation of the genetic control of migration in monarchs.

## Persistance of Migration in Monarchs and Beyond

Our work suggests that even if migration goes fully "quasi-extinct" in monarchs, it may not be lost forever. Despite the loss of allelic variance during the successive bottlenecks, monarch butterflies either maintained the genetic variation needed to migrate or secondarily re-aquired it via mutation. This gives us some hope that the contemporary loss of migration that we have observed in many migrants may be reversible. At the very least, we can be assured that for monarchs in particular, migrants in Australia constitute a reservoir of migratory alleles that could be tapped for North America if needed.

It is difficult to say what these results entail for the conservation of migratory species as a whole. While monarchs may be able to quickly recover from the loss of migratory behavior should conditions improve, it is unlikely that this is the case for most other migratory species of conservation concern, particularly for vertebrates. For example, if we assume the latter hypothesis (migratory variance in Australian monarchs is the result of a new mutation), it is clear that monarchs have had ample evolutionary opportunity to re-acquire migratory mutations: monarchs have a generation time of approximately seven generations per year if continuously breeding, which means that monarchs in Australia went through roughly ~ `r 7*50` generations in the 50 or so years since they were first reported on the continent, and although they likely experienced a strong bottleneck initially, they were probably at a relatively large effective size for most of that time. Taken together, monarchs would have had a large opportunity space for the generation of new migratory alleles.

Alternatively, it is safe to assume that if migratory variance in Australian monarchs was maintained from North America, it is clear that this would only have been possible because selection was not actively acting against that variance. This is actually quite likely: monarchs in the Pacific are not exposed to substantial annual changes in day length or average temperature, and thus do not receive their main migratory cues. Environmentally triggered migratory variation, therefore, is probably selectively nearly neutral [although more fixed migratory variance, such as wing morphology, was likely selected against, driving the observed, repeated decreases in wing size and length across newly non-migratory monarch populations, see @Freedman2020], and is therefore more likely to be maintained in non-migratory populations.

In contrast, take for example Chinook and steelhead salmon (*Oncorhynchus tshawytscha* and *mykiss*, respectively). Pre-mature migration, a migratory life history variant in both species where some portion of the population return to their natal streams earlier than others, seems to be controlled by a ancestrally derived (but independently evolved) allelic variant at the GREB1L gene, which has been lost from the majority of populations due to [@Princee1603198; @Thompson177]. Salmon have *much* longer generation times (several years in both species), and so we should not expect pre-mature migration to quickly re-evolve in salmon as did migration in monarchs, and the pre-mature migration phenotypes have been selected against due to dam construction and other anthropogenic influences [@Thompson177], meaning that GREB1L is unlikely to be selectively netural in most populations.

Nonetheless, our findings are still optimistic for the future of migration in monarch butterflies or wherever selection on migratory alleles is merely relaxed instead of shifting to be increasingly negative.

# Conclusion

This study suggests that reproductive diapause in Australian monarch butterflies is influenced by a novel genetic mechanism via the DPOGS208560 gene, and that the previously identified migratory genetic elements, such as Collagen IV $\alpha$-1, were not associated with diapause in those populations. More research is needed to determine if the migratory variant of this gene is the result of ancestral variation maintained from North America or a new mutation. Regardless, our work is consistent with the hypothesis that proximate genetic mechanisms for migration are more likely to quickly re-emerge during a transition to a migratory life history than facultative mechanisms.

# Acknowledgments

# References
<div id="refs"></div>

# Tables and Figures

```{r Table1,echo=FALSE}
Tab1 <- table(sample.meta(dat)[,c("year", "likely_pheno")])
Tab1 <- addmargins(Tab1)
Tab1 <- reshape2::dcast(as.data.frame(Tab1), year ~ likely_pheno, value.var = "Freq")
Tab1$year <- as.character(Tab1$year)
Tab1[Tab1 == "Sum"] <- "Total"
colnames(Tab1) <- c("Year", "# In Diapause", "# Reproductive", "# Total")
# Tab1 <- gt::gt(Tab1, caption = c("Table 1: Number of individuals in diapuase or that are reproductive for each year.")) %>%
#   gt::cols_label(year = "Year", M = gt::html("N<sub>subscript</sub>"), R = gt::html("N<sub>Reproductive</sub>"), Sum = gt::html("N<sub>Total</sub>"))

# kableExtra::kable(Tab1, format = "latex", booktabs = T, linesep ="\\addlinespace", caption = c("Table 1: Number of individuals in diapuase or that are reproductive for each year."), col.names = c("Year", "Diapause", "Reproductive", "Total")) %>% kableExtra::kable_styling()

# kableExtra::kable(Tab1,
#              caption = "Table 1: Contingency Table for the number of diapaused and reproductive individuals for each year.",
#              col.names = c("Year", "# Diapause", "# Reproductive", "# Total"), escape = FALSE) %>% kableExtra::kable_styling()

flextable::regulartable(Tab1, cwidth = 1.2) %>% flextable::set_caption(., "Counts of individuals per reproductive status in each year.")
```

```{r Figure_1_print,echo=FALSE,warning=FALSE,fig.cap="Figure 1: Principal Component Analysis of the monarchs from 2016 and 2018, colored by maternal family.",fig.width=8, fig.height=8, fig.align='center', fig.cap.style="Image Caption"}

pca$data$pca$MotherID <- factor(pca$data$pca$MotherID, levels = sort(as.numeric(unique(pca$data$pca$MotherID))))
colnames(pca$data$pca)[which(colnames(pca$data$pca) == "year")] <- "Year"

ggplot(pca$data$pca, aes(x = PC1, y = PC2, shape = Year, color = MotherID)) + geom_point(size = 4) +
  theme_bw() + scale_color_viridis_d(name = "Maternal Family") + ylab(pca$plots$pca$labels$y) + xlab(pca$plots$pca$labels$x)
```

```{r Figure_2_print,echo=FALSE,warning=FALSE,fig.cap="Figure 2: Percentage of individuals in reproductive diapuase in each maternal family (A) and  GMMAT p-values for association between each SNP and diapause status across all individuals (B). SNPs with a p-value lower that 5e-5 are highlighted in red.",fig.width=8, fig.height=6, fig.align='center', fig.cap.style="Image Caption"}
grid.arrange(Figure2a + ggtitle("A"), Figure2b + ggtitle("B"), Figure2al, layout_matrix = matrix(c(1, 2, 3, NA), nrow = 2), widths = c(1, .1))

```