### 2025-03-13 paper/Statistical_analysis.Rmd
changes in prep for resubmission. Still needs cleanup

---
title: "Statistical analysis for: Novel genetic association with migratory diapause in Australian monarch butterflies"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  word_document:
    reference_docx: "paper-files/style_ref.docx"
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    fig_width: 7
    fig_height: 6
bibliography: "paper-files/citations.bib"
csl: "paper-files/vancouver.csl"
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}\doublespacing
  - \usepackage{sectsty}\sectionfont{\fontsize{12}{12}\selectfont}
  - \usepackage{sectsty}\subsectionfont{\normalfont\itshape\fontsize{12}{12}\selectfont}
  - \usepackage[round]{natbib}
indent: true
sansfont: Times New Roman
fontsize: 12pt
always_allow_html: true
---

# Part 1: Map and manhattan
Setup:
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(snpR); library(data.table); library(ggplot2); library(dplyr);
library(cowplot); library(openxlsx); library(nlme); library(gggenes);
library(Biostrings)

library(rnaturalearth)
library(ggplot2)
library(sf)
library(ggspatial)
library(ggimage)

```

Figure 1 (needs edited a bit later in some other software to produce final figure).
```{r base_map, include=FALSE}
world <- ne_countries(scale = 'small', returnclass = 'sf')

polygon <- st_polygon(x = list(rbind(c(-0.0001, 90),
                                     c(0, 90),
                                     c(0, -90),
                                     c(-0.0001, -90),
                                     c(-0.0001, 90)))) |>
  st_sfc() |>
  st_set_crs(4326)

sf::sf_use_s2(FALSE)

world <- world |> st_difference(polygon)

world_robin <- st_transform(world, crs = '+proj=robin +lon_0=180 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs')

sf::sf_use_s2(TRUE)

# world <- ggplot() +
#   geom_sf(data = world_robin, fill= "azure2")  + #background color for land
#   coord_sf(xlim = c(-70e5, 100e5),
#            ylim = c(-60e5, 58e5)) +
#   theme_bw() +
#   annotation_scale(text_cex = 1.5) + 
#   theme(axis.text = element_text(size = 20),
#         axis.title = element_text(size = 24)) +
#   annotation_north_arrow(style = north_arrow_fancy_orienteering(text_size = 20), location = "bl", 
#                          pad_y = grid::unit(.05, "native"))
# 
# world
# ggsave("results/map.jpg", world, "jpg", width = 15, height = 11) # note, need to annotate this with ranges and arrows to produce final image

world <- ggplot() +
  geom_sf(data = world_robin, fill= "#dec079")  + #background color for land
  coord_sf(xlim = c(-70e5, 100e5),
           ylim = c(-60e5, 58e5)) +
  theme_bw() +
  annotation_scale(text_cex = 1.5, bar_cols = c("#eb612e", "#dec079"), text_family = "serif", text_col = "#393e42", pad_y = unit(0.75, "cm")) +
  theme(axis.text = element_text(size = 20, color = "#393e42"),
        axis.title = element_text(size = 24),
        panel.grid = element_line(color = "#393e42")) +
  annotation_north_arrow(style = north_arrow_fancy_orienteering(text_size = 20,
                                                                text_col = "#393e42",
                                                                text_family = "serif",
                                                                fill = c("#eb612e", "#dec079"),
                                                                line_col = "#393e42"),
                         location = "bl",
                         pad_y = grid::unit(.05, "native"))

world
ggsave("Figure1_base.jpg", world, "jpg", width = 15, height = 11)
```

Read in the metadata:
<!-- chunks located here to ensure everything calls correctly -->
```{r, read_sample_meta_get_numbers, echo=FALSE,message=FALSE,include=FALSE,warning=FALSE}
sample.meta <- read.table("../data/sample_metadata.txt", header = T, stringsAsFactors = F)
old.data <- data.table::fread("../data/Aus_2018_metadata.txt", header = T)
old.data[,fam := as.numeric(gsub("_.+", "", ID))]

meta.micah <- data.table::fread("../data/2016_mon_meta.txt", header = TRUE)
```

Produce figure 2, including reading in the association test results:

```{r import_and_diversity, echo=FALSE, message=FALSE, warning=FALSE,include=FALSE}
#==========prep data==========
# read in genotypes and metadata
sample.meta <- read.table("../data/sample_metadata_v4.txt", header = T, stringsAsFactors = F)

# remove samples with no phenotypic data
# nas <- which(is.na(sample.meta$likely_pheno))
# genos <- genos[,-which(is.na(sample.meta$likely_pheno))]
# sample.meta <- sample.meta[-which(is.na(sample.meta$relaxed_pheno)),]

# add in colony info
fam_info <- sample.meta(readRDS("../colony/colony_R_outputs.RDS")$col_res$x)
fam_info$MotherID <- as.numeric(gsub("#", "", fam_info$MotherID))
fam_info$MotherID <- factor(fam_info$MotherID, levels = 1:max(fam_info$MotherID))
sample.meta$MotherID <- fam_info$MotherID[match(sample.meta$SampleID, fam_info$SampleID)] 
fam_info$FatherID <- as.numeric(gsub("\\*", "", fam_info$FatherID))
fam_info$FatherID <- factor(fam_info$FatherID, levels = 1:max(fam_info$FatherID))
sample.meta$FatherID <- fam_info$FatherID[match(sample.meta$SampleID, fam_info$SampleID)] 
sample.meta$cluster <- fam_info$ClusterIndex[match(sample.meta$SampleID, fam_info$SampleID)]

# send year to factor, import, and filter
sample.meta$year <- as.factor(sample.meta$year)

#==========PCA plot==============
PCA <- fread("../scripts/association/PCAngsd_out.cov")
e <- eigen(as.matrix(PCA))
PCA <- e$vectors
PCA <- as.data.table(PCA)
colnames(PCA) <- paste0("PC", 1:ncol(PCA))
t.sizes <- c(18, 12)
FigureS1 <- ggplot(PCA, aes(x = PC1, y = PC2)) +
  geom_point(size = 3) +
  theme_bw() + 
  theme(legend.text = element_text(size = t.sizes[2]),
        axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        axis.text.y = element_text(size = t.sizes[2]),
        legend.title = element_text(size = t.sizes[1])) +
  xlab(paste0("PC1 (", round(e$values[1]/sum(e$values)*100, 3), "%)")) +
  ylab(paste0("PC2 (", round(e$values[2]/sum(e$values)*100, 3), "%)"))

```

```{r Figure2, echo=FALSE, warning=FALSE, message=FALSE}
#==========family bias===================
bias <- sample.meta[,c("MotherID", "FatherID", "quant_pheno")]
bias$quant_pheno <- bias$quant_pheno/2
bias$MotherID <- as.factor(bias$MotherID)
bias$FatherID <- as.factor(bias$FatherID)

# permute to show expected random range per rank
perms <- matrix(0, 1000, length(unique(bias$MotherID)))
for(i in 1:nrow(perms)){
  tb <- bias
  tb$quant_pheno <- sample(tb$quant_pheno, nrow(tb), TRUE)
  perm_p <- tapply(tb$quant_pheno, tb$MotherID, mean)
  perm_p <- as.data.frame(perm_p)
  perm_p$MotherID <- 1:nrow(perm_p)
  colnames(perm_p)[1] <- "score"
  perm_p$rank <- rank(perm_p$score, ties.method = "first")
  perm_p$total <- as.numeric(table(tb$MotherID))
  perms[i,] <- perm_p$score[order(perm_p$rank)]
}
perm_summary <- data.table(score = colMeans(perms),
                           rank = 1:ncol(perms))
perm_summary[, c("lCI", "uCI") := as.data.frame(t(apply(perms, MARGIN = 2, FUN = function(x){
  return(t.test(x)$conf.int)
})))]

# anova to show significance
m1 <- lm(quant_pheno ~ MotherID, bias)
m1_an <- anova(m1)

# convert for plotting
bias_p <- tapply(bias$quant_pheno, bias$MotherID, mean)
bias_p <- as.data.frame(bias_p)
bias_p$MotherID <- 1:nrow(bias_p)
colnames(bias_p)[1] <- "score"
bias_p$rank <- rank(bias_p$score, ties.method = "first")
bias_p$total <- as.numeric(table(bias$MotherID))

# generate plot
t.sizes <- c(18, 12)
Figure2a <- ggplot(bias_p, aes(x = rank, y = score, size = total)) +
  geom_point(color ="#393e42") + theme_bw() +
  geom_line(data = perm_summary, aes(x = rank, y = score), inherit.aes = FALSE, color = "#eb612e") +
  geom_ribbon(data = perm_summary, aes(ymin = lCI, ymax = uCI, x = rank), 
              inherit.aes = FALSE, color = "#eb612e", alpha = .2) +
  xlab("Maternal Family") +
  ylab("Mean Reproductive Score") +
  labs(size = "Family Size")  +
  theme(legend.text = element_text(size = t.sizes[2] - 2),
        axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        axis.text.y = element_text(size = t.sizes[2]),
        legend.title = element_text(size = t.sizes[2]),
        title = element_text(size = t.sizes[1] + 2)) +
  scale_size_continuous(range = c(5, 10)) +
  ggtitle("B")

# ggsave("~/2022-2023/ICCB2023/scatter.jpg", Figure3a, device = "jpg", dpi = 1000, height = 4, width = 8, scale = 3)

#==========association test -- read in from ANGSD============
chr_key <- fread("../data/chr_info.tsv")
chr_key$Chr_ID <- chr_key$`Chromosome name`
chr_key$Chr_ID <- make.unique(chr_key$Chr_ID)
chr_key$partID <- as.numeric(gsub(".+\\.", "", chr_key$Chr_ID))
chr_key$partID[-grep("\\.", chr_key$Chr_ID)] <- 0
chr_key <- dplyr::arrange(chr_key, as.numeric(`Chromosome name`), partID)
adat_quant <- data.table::fread("../scripts/association/association_out_quant_cov.lrt0.gz", header = T)
adat_quant$chr <- chr_key$Chr_ID[match(adat_quant$Chromosome, chr_key$`RefSeq seq accession`)]
adat_quant[grep("Un", chr), chr := "Other"]
adat_quant$pval <- pchisq(adat_quant$LRT, 1, lower.tail = FALSE)
adat_quant$lpval <- -log(adat_quant$pval)
adat_quant$fdr <- p.adjust(adat_quant$pval, "fdr")
adat_quant <- dplyr::arrange(adat_quant, pval)

q_quant <- plot_qq(adat_quant, "pval")
Figure2b <- ggplot(q_quant$.base$data, aes(x = .e, y = .o)) +
  geom_point(color = "#393e42") +
  geom_abline(slope = 1, intercept = 0, color = "#eb612e") +
  theme_bw() +
  theme(legend.text = element_text(size = t.sizes[2]),
        axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        axis.text.y = element_text(size = t.sizes[2]),
        legend.title = element_text(size = t.sizes[1]),
        title = element_text(size = t.sizes[1] + 2)) +
  ylab(bquote(Observed ~ log[10](p))) +
  xlab(bquote(Expected ~ log[10](p))) +
  ggtitle("C")

# note -- color scheme for posters!
ord <- unique(chr_key$Chr_ID, sort = F)
ord <- ord[-grep("Un", ord)]
ord <- ord[-which(ord == "Z")]
ord <- c("Z", ord, "Other")
Figure2c <- plot_manhattan(adat_quant, "fdr", chr = "chr", bp = "Position",
                           significant = 0.05, suggestive = .1, log.p = TRUE,
                           chr_order = ord,
                           highlight = FALSE, colors = c("#eb612e", "#393e42"), 
                           t.sizes = c(t.sizes[1], t.sizes[1], t.sizes[2]), 
                           simplify_output = TRUE) # highlight top snp

# refetch chr centers, pulled from snpR
chr.info <- tapply(Figure2c$data[,"cum.bp"], Figure2c$data[,"chr"], max) # chromosome lengths
chr.centers <- rowMeans(cbind(tapply(Figure2c$data[,"cum.bp"], Figure2c$data[,"chr"], min), chr.info)) # chromosome centers

# plot
Figure2c <- Figure2c +
  ylab("-log10(p-value)") +
  xlab("Chromosome") +
  theme(legend.text = element_text(size = t.sizes[2]),
        axis.text = element_text(size = t.sizes[2]),
        axis.title = element_text(size = t.sizes[1]),
        axis.text.y = element_text(size = t.sizes[2]),
        legend.title = element_text(size = t.sizes[1]),
        title = element_text(size = t.sizes[1] + 2)) +
  scale_x_continuous(expand = c(0, 0), breaks = chr.centers, labels = 
                       names(chr.centers)) +
  ggtitle("D")

# ggsave("~/2022-2023/ICCB2023/manhattan.jpg", Figure3b, device = "jpg", dpi = 1000, height = 4, width = 8, scale = 3)

#==========egg photos (actual A)=========
imgs <- data.frame(category = c("Mature", "Partially Mature", "Immature"),
                   img = c("~/GitHub/aus_monarchs/paper/paper-files/mature.JPG",
                           "~/GitHub/aus_monarchs/paper/paper-files/partially_mature.JPG",
                           "~/GitHub/aus_monarchs/paper/paper-files/immature.JPG"),
                   x = 1,
                   y = 1)
imgs$category <- factor(imgs$category, levels = c("Mature", "Partially Mature", "Immature"))

eggplot <- ggplot(imgs, aes(x = x, y = y, image = img)) +
  geom_image(size = 1.3) +
  facet_wrap(~category, ncol = 1, strip.position = "left") +
  theme_minimal() +
  theme(strip.text = element_text(size = t.sizes[2], angle = 90),
        strip.placement = "outside",
        title = element_text(size = t.sizes[1] + 2),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank()) +
  ggtitle("A")

#==========combine=======================
f2al <- ggpubr::get_legend(Figure2a)
Figure2_top <- plot_grid(eggplot, Figure2a + guides(size = "none"), 
                         f2al, Figure2b, 
                         align = "hv", axis = "tlbr",
                         nrow = 1,
                         rel_widths = c(.5, 1, .2, 1)) # note, A is actually egg pictures, these labels are off

Figure2 <- plot_grid(Figure2_top,
                     Figure2c,
                     nrow = 2)

save_plot("Figure2.pdf", Figure2, base_height = 11, base_width = 15)

```

Re-analyze the old wing morphology data to show correlated migratory syndrome.
```{r re_analyze, include=FALSE}
old_phenotypic <- read.xlsx("../data/monarch_development.xlsx")

mod_AsRa <- lme(Aspect_Ratio ~ Yolked_Oocytes, random = ~ 1|Female_Line, old_phenotypic, na.action = na.omit)
mod_Area <- lme(Marea ~ Yolked_Oocytes, random = ~ 1|Female_Line, old_phenotypic, na.action = na.omit)
```

# Analyize genes of interest
Produce Figure 3:
```{r genes_of_interest}
#=======identify targets=========
window <- 50*1000 # 10kb
targets <- adat_quant[fdr <= 0.05,]
targets[,start := Position - window]
targets[,end := Position + window]
targets <- dplyr::arrange(targets, Chromosome, start)

# one peak per chr, so simplify
targets[,start := min(start), by = chr]
targets[,end := max(end), by = chr]
targets <- unique(targets[,c("Chromosome", "chr", "start", "end")])

#=======locate genes=============
gtf <- rtracklayer::readGFF("../data/genomic.gtf")
gtf <- as.data.table(gtf)
gtf[,direction := ifelse(strand == "+", TRUE, FALSE)]
keep <- vector("list", nrow(targets))
all_genes <- keep
for(i in 1:nrow(targets)){
  keep[[i]] <- gtf[which(seqid == targets[i,Chromosome] &
                     ((end >= targets[i,start] & end <= targets[i,end]) |
                        (start <= targets[i,end] & start >= targets[i,start]) |
                        (start <= targets[i, start] & end >= targets[i, end]))),]
  all_genes[[i]] <- unique(keep[[i]][type == "gene",])
}
all_genes <- rbindlist(all_genes, idcol = "target")
IDs <- gsub("GeneID:", "", all_genes$db_xref)

# get the prot sequences from NCBI
cmd <- paste0("~/bin/datasets download gene gene-id ", paste0(IDs, collapse = " "))
system(cmd)

# copy the .faa and paste into monarchbase, read in the result here
prot_matches <- fread("../data/prot_matches_monarchbase.txt")
prot_matches[,LOC := unlist(stringi::stri_extract_all_regex(`Query ID`, "LOC.{9}"))]
prot_matches <- dplyr::arrange(prot_matches, LOC, Evalue)
prot_matches <- prot_matches[, head(.SD, 1), by= `Query ID`]
prot_matches <- dplyr::arrange(prot_matches, Length)
prot_matches <- prot_matches[, head(.SD, 1), by= LOC]

all_genes <- merge(all_genes, prot_matches, by.x = "gene_id", by.y = "LOC", all = TRUE)

#========plot====================
plot_parts <- vector("list", nrow(targets))
pmax <- max(-log10(adat_quant$fdr))

for(i in 1:nrow(targets)){
  # figure out max and min range
  min_pos <- targets[i,start]
  max_pos <- targets[i,end]

  tgtf <- keep[[i]]
  subgene <- tgtf[type == "CDS",]
  tgtf <- tgtf[type == "gene",]
  subgene[,gene_start := tgtf$start[match(gene, tgtf$gene)]]
  subgene[,gene_end := tgtf$end[match(gene, tgtf$gene)]]

  tgtf[,gene_lab := start + ((end - start)/2)]
  tgtf$OGS_name <- all_genes$`Monarch geneid`[match(tgtf$gene_id, all_genes$gene_id)]
  tgtf$OGS_name[is.na(tgtf$OGS_name)] <- tgtf$gene_id[is.na(tgtf$OGS_name)]

  subgene[,OGS_name := tgtf$OGS_name[match(gene, tgtf$gene)]]

  if(min_pos > min(c(tgtf$start, tgtf$end))){
    min_pos <- min(c(tgtf$start, tgtf$end))
  }
  if(max_pos < max(c(tgtf$start, tgtf$end))){
    max_pos <- max(c(tgtf$start, tgtf$end))
  }

  # plot SNPs

  tsnps <- adat_quant[which(Chromosome == targets[i,Chromosome] &
                     Position >= min_pos & Position <= max_pos),]

  pt <- ggplot(tsnps, aes(x = Position/1000000, y = -log10(fdr))) +
    geom_point() +
    theme_bw() +
    scale_x_continuous(limits = c(min_pos/1000000, max_pos/1000000)) +
    scale_y_continuous(expand = c(0, .05), limits = c(0, pmax))

  pmax <- max(c(pmax, max(-log10(tsnps$fdr))))

  if(i == 1)
    pt <- pt + 
    theme(axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.text.y = element_text(size = t.sizes[2]),
          axis.title.y = element_text(size = t.sizes[1]),
          axis.ticks.x = element_blank()) +
  ylab(bquote(-log[10](p[FDR])))

  if(i != 1){
    pt <- pt + 
      theme(axis.text.x = element_blank(),
          axis.title = element_blank(),
          axis.text.y = element_text(size = t.sizes[2]),
          axis.ticks.x = element_blank())
  }

  tgtf$OGS_name[tgtf$OGS_name == "DPOGS204613"] <- paste0("Karst/DPOGS204613")
  subgene$OGS_name[subgene$OGS_name == "DPOGS204613"] <- paste0("Karst/DPOGS204613")

  pb <- ggplot(tgtf, aes(xmin = start/1000000, xmax = end/1000000, 
                         y = OGS_name, forward = direction, fill = gene)) +
    geom_gene_arrow(arrow_body_height = unit(1, "lines"),
                    arrowhead_height = unit(1.5, "lines"), show.legend = FALSE,
                    fill = "#E5C16F") + 
    geom_subgene_arrow(data = subgene, 
                       aes(xmin = gene_start/1000000, xmax = gene_end/1000000, 
                           xsubmin = start/1000000, xsubmax = end/1000000,
                           y = OGS_name), 
                       fill = "black", arrow_body_height = unit(1, "lines"),
                       arrowhead_height = unit(1.5, "lines")) +
    theme_bw() +
    scale_x_continuous(limits = c(min_pos/1000000, max_pos/1000000)) +
    ylab("Gene") +
    xlab(paste0("Chromosome ", targets[i]$chr, " Position (mb)"))

  if(i == 1)
    pb <- pb + 
    theme(axis.text = element_text(size = t.sizes[2]),
          axis.title = element_text(size = t.sizes[1]))

  if(i != 1){
    pb <- pb + 
      theme(axis.title.y = element_blank(),
          axis.text = element_text(size = t.sizes[2]),
          axis.title.x = element_text(size = t.sizes[1]))
  }

  plot_parts[[i]] <- plot_grid(pt, pb, ncol = 1, rel_heights = c(1, 1),
                               align = "hv", axis = "tl")

}

Figure3 <- plot_grid(plot_parts[[1]],
                     plot_parts[[2]] + theme(axis.title.y = element_blank()),
                     plot_parts[[3]] + theme(axis.title.y = element_blank()), nrow = 1) +
  theme(plot.margin = unit(c(.1, .5, .1, .5), "cm"))

save_plot("Figure3.pdf", Figure3, base_height = 11, base_width = 15)

```

# Misc statistics for paper
Mapping stats:
```{r mapping_stats}
sort_flagstat <- list.files("../data/v4/flagstat/", "\\.sort\\.flagstat", full.names = TRUE)
flt_flagstat <- list.files("../data/v4/flagstat/", "\\.sort\\.flt\\.flagstat", full.names = TRUE)

sort_flagstat <- unlist(lapply(sort_flagstat, readLines))
flt_flagstat <- unlist(lapply(flt_flagstat, readLines))

mapping <- data.table(sample = sample.meta$SampleID,
                      reads = as.numeric(gsub(" .+", "", sort_flagstat[seq(1, length(sort_flagstat), by = 16)])),
                      mapped = as.numeric(gsub(" .+", "", sort_flagstat[seq(7, length(sort_flagstat), by = 16)])),
                      retained = as.numeric(gsub(" .+", "", flt_flagstat[seq(1, length(sort_flagstat), by = 16)])))

total_reads <- sum(mapping$reads)
total_mapped <- sum(mapping$mapped)
total_retained <- sum(mapping$retained)

percent_mapped <- total_mapped/total_reads
percent_retained <- total_retained/total_reads

```

Table S1
```{r Table_S1}

TS1 <- all_genes[,c("gene_id", "Monarch geneid", "seqid", "start", "end")]
TS1$Chromosome <- targets$chr[match(TS1$seqid, targets$Chromosome)]
colnames(TS1) <- c("RefSeq ID", "OGS ID", "Chromosome RefSeq ID", "Start", "End", "Chromosome")
TS1 <- TS1[,c("RefSeq ID", "OGS ID", "Chromosome", "Chromosome RefSeq ID", "Start", "End")]
TS1 <- dplyr::arrange(TS1, Chromosome, Start)
fwrite(TS1, "TableS1.txt", sep = "\t", col.names = TRUE, row.names = FALSE, quote = FALSE)

```

Table 1:
```{r Table1,echo=FALSE}
Tab1 <- table(sample.meta[,c("year", "quant_pheno")])
Tab1 <- addmargins(Tab1)
Tab1 <- reshape2::dcast(as.data.frame(Tab1), year ~ quant_pheno, value.var = "Freq")
Tab1$year <- as.character(Tab1$year)
Tab1[Tab1 == "Sum"] <- "Total"
colnames(Tab1) <- c("Year", "# Immature", "# Partially Mature", "# Mature", "Total")
Tab1$`# Maternal Familes` <- c(dplyr::arrange(as.data.table(sample.meta[,c("year", "MotherID")])[,list(nfam = length(unique(MotherID))), by = year], year)$nfam,
                               length(unique(sample.meta$MotherID)))
# Tab1 <- gt::gt(Tab1, caption = c("Table 1: Number of individuals in diapuase or that are reproductive for each year.")) %>%
#   gt::cols_label(year = "Year", M = gt::html("N<sub>subscript</sub>"), R = gt::html("N<sub>Reproductive</sub>"), Sum = gt::html("N<sub>Total</sub>"))

# kableExtra::kable(Tab1, format = "latex", booktabs = T, linesep ="\\addlinespace", caption = c("Table 1: Number of individuals in diapuase or that are reproductive for each year."), col.names = c("Year", "Diapause", "Reproductive", "Total")) %>% kableExtra::kable_styling()

# kableExtra::kable(Tab1,
#              caption = "Table 1: Contingency Table for the number of diapaused and reproductive individuals for each year.",
#              col.names = c("Year", "# Diapause", "# Reproductive", "# Total"), escape = FALSE) %>% kableExtra::kable_styling()

flextable::regulartable(Tab1, cwidth = 1.2) %>% flextable::set_caption(., "Total count of immature, partially mature, and mature female monarch butterflies reared and sequenced in each sampling year.")
```

# Karst across the pacific and models
## Read in
Read in the gprobs and the called, phased data.

```{r}
genos <- fread("D://Cluster_storage/comparison_monarchs/candidate_region_gl_imputed.candidate_region.beagle.gz.phased.gz")

gprobs <- fread("D://Cluster_storage/comparison_monarchs/candidate_region_gl_imputed.candidate_region.beagle.gz.gprobs.gz")

bamlist <- fread("D://Cluster_storage/comparison_monarchs/bamlist.txt", header = FALSE)
```

Figure out which calls have low confidence in
```{r}
gprobs <- gprobs[,-c(1:3)]
# loop through each individual
prog <- 1
OK <- matrix(TRUE, nrow(gprobs), ncol = ncol(gprobs)/3)
for(i in 1:(ncol(gprobs)/3)){
  max_freq <- matrixStats::rowMaxs(as.matrix(gprobs[,prog:(prog + 2)]))
  OK[,i] <- ifelse(max_freq >= 0.95, TRUE, FALSE)
  prog <- prog + 2
}
```

Joint mark loci, individuals to remove below 50% called.

```{r}
mloc <- rowMeans(OK) # missing per locus
mind <- colMeans(OK) # missing per ind

bad_ind <- which(mind < .6)
bad_loc <- which(mloc < .6)

OK <- OK[-bad_loc, -bad_ind]
```

Distribution of OK imputation
```{r}
plot(colMeans(OK))
```

## Prepare metadata
```{r}
sample.meta <- read.table("../data/sample_metadata_v4.txt", header = T, stringsAsFactors = F)
matches <- numeric(nrow(sample.meta))
plates <- stringi::stri_extract(bamlist$V1, regex = "plate_[1-3]")
plates <- gsub("_", "", plates)
barcode <- stringi::stri_extract(bamlist$V1, regex = "RA_GG[A-Z]{13}\\.")
barcode <- gsub("RA_GG", "", barcode)
barcode <- gsub("TGCAG\\.", "", barcode)
# for(i in 1:length(matches)){
#   matches[i] <- intersect(which(sample.meta$Index[i] == barcode), which(sample.meta$Plate[i] == plates))
# }

meta <- data.table(barcode = bamlist$V1)
meta[is.na(plates), pop := substr(barcode, 1, 3)]
meta[is.na(plates), phenotype := as.numeric(NA)]
meta[!is.na(plates), pop := "AUS"]
meta[,Index := gsub("TGCAG\\..+", "", gsub(".+_GG", "", barcode))]
meta[,Plate := gsub("rad_split.+", "", gsub("\\.\\./aus_monarchs/", "", barcode))]
meta[,Plate := gsub("_", "", Plate)]
meta <- merge(meta, sample.meta[,c("Index", "quant_pheno", "Plate")], by = c("Index", "Plate"), all.x = TRUE, sort = FALSE)
# meta[matches, phenotype := sample.meta$quant_pheno]
meta[,pop_cat := ifelse(pop == "NAM", "North America",
                        ifelse(pop %in% c("QLD", "NSW", "AUS"), "Australia", 
                               ifelse(pop == "HAW", "Hawaii", "Pacific")))]

```

## Get genotypes and allele frequencies

Remove the poor individuals and loci
```{r}
bad_ind_indices <- rep(1:((ncol(genos) - 2)/2), each = 2)
bad_ind_indices <- which(bad_ind_indices %in% bad_ind) + 2
genos <- genos[-bad_loc,-..bad_ind_indices]
meta <- meta[-bad_ind,]
```

Find loci with low mgc and remove
```{r}
# combine, then import to snpR
combined_genos <- as.matrix(genos[,-c(1:2)])
a1 <- combined_genos[,seq(1,ncol(combined_genos), by = 2)]
a2 <- combined_genos[,seq(2,ncol(combined_genos), by = 2)]
combined_genos <- paste0(a1, a2)
combined_genos <- gsub("0", "A", combined_genos)
combined_genos <- gsub("1", "C", combined_genos)
combined_genos <- gsub("2", "G", combined_genos)
combined_genos <- gsub("3", "3", combined_genos)

combined_genos <- as.data.frame(matrix(combined_genos, nrow = nrow(genos)))

d <- import.snpR.data(combined_genos, snp.meta = genos[,1:2], sample.meta = meta)

# which have bad genotype_counts
df <- filter_snps(d, mgc = 5) # nothing removed so no need to go further with this filter. Makes sense given our 0.05 maf during calling
```

Fetch Positions
```{r}

positions <- genos$id
positions <- as.numeric(gsub(".+_", "", positions))
# karst_snps <- which(positions >= 444283  & positions <= 503399)
# genos <- genos[karst_snps,]

```

## add in colony info and prep for models
```{r}
hap_meta <- meta[rep(1:nrow(meta), each = 2),]

# add in colony info
families <- read.table("../data/sample_metadata_v4.txt", header = T, stringsAsFactors = F)
fam_info <- sample.meta(readRDS("../colony/colony_R_outputs.RDS")$col_res$x)
fam_info$MotherID <- as.numeric(gsub("#", "", fam_info$MotherID))
fam_info$MotherID <- factor(fam_info$MotherID, levels = 1:max(fam_info$MotherID))
families$MotherID <- fam_info$MotherID[match(families$SampleID, fam_info$SampleID)] 
fam_info$FatherID <- as.numeric(gsub("\\*", "", fam_info$FatherID))
fam_info$FatherID <- factor(fam_info$FatherID, levels = 1:max(fam_info$FatherID))
families$FatherID <- fam_info$FatherID[match(families$SampleID, fam_info$SampleID)] 
families$cluster <- fam_info$ClusterIndex[match(families$SampleID, fam_info$SampleID)]

# mod_data <- hap_meta[pop == "AUS",]
# mod_data[,Index := gsub("TGCAG\\..+", "", gsub(".+_GG", "", barcode))]
# colnames(mod_data)[which(colnames(mod_data) == "cluster")] <- "Haplotype"
# mod_data <- merge(mod_data, families, by = c("Index", "Plate", "quant_pheno"),  sort = FALSE)
# 
# mod1 <- lmerTest::lmer(quant_pheno ~ as.factor(Haplotype) + (1|cluster), mod_data)
# mod2 <- lmerTest::lmer(quant_pheno ~ (1|cluster), mod_data)
# summary(mod1)
# anova(mod1, mod2)

```

get only the largest effect loci, construct models
```{r}

# pull in asso data
chr_key <- fread("../data/chr_info.tsv")
chr_key$Chr_ID <- chr_key$`Chromosome name`
chr_key$Chr_ID <- make.unique(chr_key$Chr_ID)
chr_key$partID <- as.numeric(gsub(".+\\.", "", chr_key$Chr_ID))
chr_key$partID[-grep("\\.", chr_key$Chr_ID)] <- 0
chr_key <- dplyr::arrange(chr_key, as.numeric(`Chromosome name`), partID)
adat_quant <- data.table::fread("../scripts/association/association_out_quant_cov.lrt0.gz", header = T)
adat_quant$chr <- chr_key$Chr_ID[match(adat_quant$Chromosome, chr_key$`RefSeq seq accession`)]
adat_quant[grep("Un", chr), chr := "Other"]
adat_quant$pval <- pchisq(adat_quant$LRT, 1, lower.tail = FALSE)
adat_quant$lpval <- -log(adat_quant$pval)
adat_quant$fdr <- p.adjust(adat_quant$pval, "fdr")
adat_quant <- dplyr::arrange(adat_quant, pval)
adat_quant <- adat_quant[chr == 29 & fdr <= 0.05,]

# positions
top_loci <- which(positions %in% adat_quant$Position)

top_loci <- t(genos[top_loci,-c(1:2)])
tl_pos <- positions[which(positions %in% adat_quant$Position)]
top_loci <- as.data.table(cbind(hap_meta, top_loci))
## change to either 1 or 2
top_loci[,c("V1", "V2", "V3", "V4", "V5", "V6") := lapply(.SD, function(x) as.numeric(as.factor(x)) - 1), 
         .SDcols = c('V1', 'V2', 'V3', 'V4', 'V5', 'V6')]

# models
top_loci <- merge(top_loci, families, by = c("Index", "Plate", "quant_pheno"),  sort = FALSE, all.x = TRUE)
colnames(top_loci)[which(colnames(top_loci) %in% paste0("V", 1:6))] <- paste0("SNP", 1:6)
mod1L <- lmerTest::lmer(quant_pheno ~ SNP1 + SNP2 + SNP3 + SNP4 + SNP5 + SNP6 + (1|FatherID) + (1|MotherID), top_loci[pop == "AUS",]) # note, nearly perfect linkage so most vars are dropped!
mod2L <- lmerTest::lmer(quant_pheno ~ (1|FatherID) + (1|MotherID), top_loci[pop == "AUS",])
summary(mod1L)
anova(mod1L, mod2L)

# flip such that "1" is the "migratory" allele
for(i in 1:6){
  var <- paste0("SNP", i)
  form <- formula(paste0("quant_pheno ~ ", var, "+ (1|FatherID) + (1|MotherID)"))
  tm <- summary(lmerTest::lmer(form, top_loci[pop == "AUS",]))$coefficients[2,1]
  if(tm > 0){
    top_loci[[var]] <- ifelse(top_loci[,..var] == 0, 1, 0)
  }
}

# model as inds
top_loci_genos <- top_loci[seq(1, nrow(top_loci), by = 2),]
top_loci_genos <- cbind(top_loci_genos[,c(1:7, 14:24)], 
                        top_loci_genos[,.(SNP1, SNP2, SNP3, SNP4, SNP5, SNP6)] + 
                          top_loci[seq(2, nrow(top_loci), by = 2),.(SNP1, SNP2, SNP3, SNP4, SNP5, SNP6)])

mod1L <- lmerTest::lmer(quant_pheno ~ SNP1 + SNP2 + SNP3 + SNP4 + SNP5 + SNP6 + (1|FatherID) + (1|MotherID), 
                        top_loci_genos[pop == "AUS",]) # note, nearly perfect linkage so most vars are dropped!
mod2L <- lmerTest::lmer(quant_pheno ~ (1|FatherID) + (1|MotherID), top_loci_genos[pop == "AUS",])
summary(mod1L)
anova(mod1L, mod2L)
MuMIn::r.squaredGLMM(mod1L) # explains roughly 14% of the variation in diapause development

mafs <- top_loci[,lapply(.SD, mean), .SDcols =  paste0("SNP", 1:6), by = pop]
mafsm <- data.table::melt(mafs, id.vars = "pop", variable.name = "SNP", value.name = "Frequency")
# top_locim <- data.table::melt(top_loci, id.vars = colnames(top_loci)[-which(colnames(top_loci) %in% paste0("SNP", 1:6))])

mafsm[,pop_cat := ifelse(pop == "NAM", "North America",
                        ifelse(pop %in% c("QLD", "NSW", "AUS"), "Australia", 
                               ifelse(pop == "HAW", "Hawaii", "Pacific")))]

maf_plot <- ggplot(mafsm, aes(x = pop_cat, y = Frequency, color = pop_cat)) + 
  geom_jitter(height = 0, width = .1) +
  facet_wrap(~SNP) +
  theme_bw() +
  khroma::scale_color_bright()
maf_plot # note, strongest asso is SNP1

```

## plot top SNPs on a map
Producing Figure 4
```{r}
world <- ne_countries(scale = 'small', returnclass = 'sf')

polygon <- st_polygon(x = list(rbind(c(-0.0001, 90),
                                     c(0, 90),
                                     c(0, -90),
                                     c(-0.0001, -90),
                                     c(-0.0001, 90)))) |>
  st_sfc() |>
  st_set_crs(4326)

sf::sf_use_s2(FALSE)

world <- world |> st_difference(polygon)

world_robin <- st_transform(world, crs = '+proj=robin +lon_0=180 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs')

sf::sf_use_s2(TRUE)

# prep the points for plotting
NAM <- geosphere::geomean(matrix(c(c(-100.289503, 19.556050), c(-121.930681, 36.625980)), 2, 2, byrow = TRUE))
lat_long <- list(NAM = rev(NAM),
                 HAW = c(20.9997, -155.6581),
                 GUA = c(11.421207, 142.736584),
                 ROT = c(14.154628, 145.191535), SAI = c(17.201243, 147.750705),
                 SAM = c(-13.613179, -172.351278), FIJ = c(-17.924768, 178.081698),
                 NCA = c(-21.299579, 165.383757), NOR = c(-29.024356, 167.945279),
                 NSW = c(-32.751896, 151.667208),
                 VIC = c(-34.574338, 138.689131), NZL = c(-37.124496, 174.961893),
                 AUS = c(-27.540750, 152.906306))
lat_long <- t(as.data.frame(lat_long))
lat_long <- as.data.frame(lat_long)
colnames(lat_long) <- c("lat", "long")
lat_long$pop <- rownames(lat_long)
lat_long <- sf::st_as_sf(lat_long, coords = c("long", "lat"))
st_crs(lat_long) <- st_crs(rnaturalearth::ne_countries(scale = 'small', returnclass = 'sf'))
lat_long <- st_transform(lat_long, st_crs(world_robin))

cols <- c(paste0("SNP", 1:6), "pop")
# mafsm_map <- data.table::melt(top_loci[,..cols], id.vars = "pop", variable.name = "SNP", value.name = "GT")
# mafsm_map <- mafsm_map[,list(mean_frequency = mean(GT)), by = pop]
# mafsm_map <- merge(lat_long, mafsm_map, by = "pop")
mafsm_map <- data.table::melt(top_loci[,..cols], id.vars = "pop", variable.name = "SNP", value.name = "GT")
mafsm_map <- mafsm_map[SNP == "SNP1",list(`Top SNP Frequency` = mean(GT)), by = pop]
mafsm_map <- merge(lat_long, mafsm_map, by = "pop")

# mafsm_map_means <- as.data.table(mafsm_map)
# mafsm_map_means[,pop_cat := ifelse(pop == "NAM", "North America",
#                         ifelse(pop %in% c("QLD", "NSW", "AUS"), "Australia", 
#                                ifelse(pop == "HAW", "Hawaii", "Pacific")))]
# inset <- ggplot(mafsm_map_means, aes(x = pop_cat, y = mean_frequency)) +
#   geom_jitter(height = 0, width = .1, size = 3) +
#   theme_bw() +
#   khroma::scale_color_bright() +
#   xlab("Area") +
#   ylab("Mean Frequency")

mafsm[,Area := ifelse(pop == "NAM", "North America",
                      ifelse(pop %in% c("QLD", "NSW", "AUS"), "Australia", 
                             ifelse(pop == "HAW", "Hawaii", "Other Pacific")))]

tsizes <- c(14, 16, 22)
PanelB <- ggplot(mafsm, aes(x = Area, y = Frequency, color = Area)) + 
  geom_jitter(height = 0, width = .1, size = 4) +
  facet_grid(~SNP,) +
  theme_bw() +
  scale_color_manual(values = c("#eb612e", "#dec079", "#393e42", "#809a54")) +
  xlab("Area") + ylab("Diapause Allele Frequency") +
  theme(axis.text = element_text(size = tsizes[1]),
        axis.title = element_text(size = tsizes[2]),
        axis.text.x = element_blank(),
        strip.text = element_text(size = tsizes[1]), 
        strip.background = element_blank(),
        legend.title = element_text(size = tsizes[2]),
        legend.text = element_text(size = tsizes[1]),
        axis.ticks.x = element_blank(),
        plot.title = element_text(size = tsizes[3])) +
  ggtitle("B")

PanelA <- ggplot() +
  geom_sf(data = world_robin, fill = "lightgray")  + #background color for land
  geom_sf(data = mafsm_map, aes(color = `Top SNP Frequency`), size = 8) +
  # geom_sf(data = mafsm_map[mafsm_map$mean_frequency == 0,], size = 4, color = "black") +
  geom_sf(data = mafsm_map[mafsm_map$`Top SNP Frequency` == 0,], size = 8, color = "darkred", shape = 13) +
  khroma::scale_color_lajolla() + 
  coord_sf(xlim = c(-70e5, 100e5),
           ylim = c(-60e5, 58e5)) +
  theme_bw() +
  annotation_scale(text_cex = 1.5, bar_cols = c("#eb612e", "#dec079"), text_family = "serif", text_col = "#393e42", pad_y = unit(0.5, "cm")) +
  theme(axis.text = element_text(size = tsizes[1], color = "#393e42"),
        axis.title = element_text(size = tsizes[2]),
        panel.grid = element_line(color = "gray"),
        legend.title = element_text(size = tsizes[2]),
        legend.text = element_text(size = tsizes[1] - 6),
        plot.title = element_text(size = tsizes[3])) +
  annotation_north_arrow(style = north_arrow_fancy_orienteering(text_size = tsizes[1],
                                                                text_col = "#393e42",
                                                                text_family = "serif",
                                                                fill = c("#eb612e", "#dec079"),
                                                                line_col = "#393e42"),
                         location = "bl",
                         pad_y = grid::unit(.06, "native")) +
  ggtitle("A")

Figure4 <- plot_grid(PanelA, PanelB, ncol = 1, align = "vh", axis = "lr", rel_heights = c(1, .5))
save_plot("../paper/Figure4.pdf", Figure4, base_width = 13, base_height = 11)

```